---
title: "Weather Events Damage and Health Effects"
output: html_document
---

## Synopsis

The U.S. National Oceanic and Atmospheric Administration's (NOAA) collects data
on all major weather events in the US from 1950 until today. This analysis 
investigates the most harmful types of weather events across four categories:

* Fatalities
* Injuries
* Property Damage
* Crop Damage

## Data Processing

We will be using U.S. National Oceanic and Atmospheric Administration's (NOAA) 
storm database. This database tracks characteristics of major storms and weather 
events in the United States, including when and where they occur, as well as 
estimates of any fatalities, injuries, and property damage.

First download the storm data and store as a csv in your working directory. Then
read it into R and delete the csv.

```{r download_storm_data, cache=T}
download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2","storm_data.csv","curl")
storms <- read.csv("storm_data.csv")
unlink("storm_data.csv")
```

Collect some summary stats on the dataset & analyze the missign values.
```{r summary stats}
nrow = nrow(storms)
num_events <- length(unique(storms$EVTYPE))
mean_na <- mean(is.na(storms[,c("INJURIES","FATALITIES","PROPDMG","CROPDMG")]))
```

We know that there are `r nrow` observations in this dataset and 
`r num_events` different types of events to be looked at. We also know that 
of the fiels where we will be looking, there are `r mean_na` missing values,
so we can feel safe in analyzing the entire set.

Now we process the data, taking the mean number of fatalities, injuries, 
property damange, and crop damage at each.

```{r data processing}
damage_by_type <- 
  aggregate(storms[,c("INJURIES","FATALITIES","PROPDMG","CROPDMG")], 
            by=list(storms$EVTYPE), FUN="mean")
names(damage_by_type)[1] <- "EVTYPE"
```

Collect the number of events that caused damage in each category and the event
that caused the most damage in each category. Loop through the fields to find 
the index of the max and then collect the category and max for that damage type.

```{r data summary}
num_injury_events <- nrow(subset(damage_by_type, INJURIES > 0))
num_fatality_events <- nrow(subset(damage_by_type, FATALITIES > 0))
num_propdmg_events <- nrow(subset(damage_by_type, PROPDMG > 0))
num_cropdmg_events <- nrow(subset(damage_by_type, PROPDMG > 0))

injuries <- damage_by_type[,1:2]
injuries <- injuries[rev(order(injuries$INJURIES)),]
injuries <- head(injuries,5)

fatalities <- damage_by_type[,c(1,3)]
fatalities <- fatalities[rev(order(fatalities$FATALITIES)),]
fatalities <- head(fatalities,5)

propdmg <- damage_by_type[,c(1,4)]
propdmg <- propdmg[rev(order(propdmg$PROPDMG)),]
propdmg <- head(propdmg,5)

cropdmg <- damage_by_type[,c(1,5)]
cropdmg <- cropdmg[rev(order(cropdmg$CROPDMG)),]
cropdmg <- head(cropdmg,5)
```

## Results

Each of the four different types of harm has differnet events associated with
damage. We can see the top five events for each type of harm plotted below.

```{r plot}
par(mfrow=c(2,2))
with(injuries, barplot(INJURIES, names.arg=EVTYPE, las=2, cex.names=.7))
with(fatalities, barplot(FATALITIES, names.arg=EVTYPE, las=2, cex.names=.7))
with(propdmg, barplot(PROPDMG, names.arg=EVTYPE, las=2, cex.names=.7))
with(cropdmg, barplot(CROPDMG, names.arg=EVTYPE, las=2, cex.names=.7))
```

The top five events for all four type of harm are

``` {r}
injuries
```

``` {r}
fatalities
```

``` {r}
propdmg
```

``` {r}
cropdmg
```
