---
title: "Statisical Inference Course Project -- Part I"
author: "Scott Shepard"
date: "October 26, 2014"
output: pdf_document
---

This is the first of two reports for the Statistical Inference Course Project.

This report explores the exponential distribution. The exponential distribution 
$p(x)={lambda^2}$ It has a mean of 1/lambda and a standard deviation of 1/lambda.

In R we can use the function rexp(n, lambda) to simulate the exponential
distribution. We take 40 observations 1000 times with lambda = 0.2.

```{r}
  lambda = 0.2
  trials <- lapply(1:1000, function(i) rexp(40, lambda))
```
We can then take the means of each of these trials
```{r}
  trial_means <- sapply(trials, mean)
```

These trial means should be centered at the theoretical mean. So we exepect the 
median of the trial_means to be about 5.

The standard deviation of these means should be about the theoretical standard
error, which is the distribution's standard deviation devided by the square 
root of the number of observations, which is this case is 40.

```{r echo=FALSE}
  theoretical_mean <- 1/lambda
  standard_error <- (1/lambda) / sqrt(40)
  median <- median(trial_means)
  sd <- sd(trial_means)
```

The theoretical mean is `r theoretical_mean` and the trial means are centered
at `r median` so that checks out. The standard error is `r standard_error` and 
standard deviation of the trial means is `r sd` so that also checks out.

The trial means are approximately normaly distributed. We can see this by looking
at a histogram of the trial means.
```{r}
  hist(trial_means, main="Distribution of Means")
```  

See, normal distribution.